Mamba vs SOTA for Vision Tasks\\ \thanks{$\dagger$ Equal contributions}

### 2024-04-28 paper/main.tex
updated with ViT ViM comparison

\begin{abstract}
    Abstract
    This paper explores the performance of the Mamba architecture, a variant of Structured State Space Sequence Models (SSMs), in comparison to state-of-the-art models such as Transformers, U-Net, and ResNet for vision tasks. We focus on image segmentation and classification, with the goal of maintaining high accuracy while keeping parameter counts low. Developing compact models with fewer parameters is crucial for the future of AI, not only to reduce energy consumption but also to enable deployment on resource-constrained edge devices.
    By leveraging the selective state spaces in Mamba blocks, we aim to achieve efficient and effective segmentation and classification performance while maintaining the linear scalability of SSMs. We implement Mamba-based architectures and compare their results to popular models like U-Net, ResNet, and Transformer-based approaches on standard vision benchmarks.
    In addition to accuracy, we place a strong emphasis on model size, targeting compact architectures suitable for resource-constrained environments. Through careful design choices and optimizations, we strive to develop Mamba-based models that achieve competitive accuracy with significantly fewer parameters compared to existing state-of-the-art models.
    This work highlights the potential of Mamba and SSMs for efficient vision tasks, contributing to the development of compact yet accurate models in image segmentation and classification. We open-source our code to facilitate further research and exploration of these architectures: \href{https://github.com/lxaw/mamba-vs-else-vision}{https://github.com/lxaw/mamba-vs-else-vision}
\end{abstract}

@@

\section{Introduction}
The rapid advancement of deep learning has led to remarkable achievements in various domains, including computer vision, natural language processing, and speech recognition. However, this progress has been accompanied by a significant increase in the size and complexity of neural network models. State-of-the-art models often have several million or even billions of parameters, making them computationally expensive and challenging to deploy on resource-constrained devices \cite{brown2020language, dosovitskiy2021image}. This trend poses concerns regarding energy consumption and the feasibility of deploying these models on edge devices, which have limited memory and processing power.

The growth of edge devices and the emerging field of TinyML \gr{TinyML? not a field but is a library?} have highlighted the need for more efficient and compact models \cite{warden2019tinyml}. Edge devices, such as smartphones, wearables, and IoT sensors, have become increasingly prevalent in our daily lives. These devices often require real-time processing and decision-making capabilities, but their limited resources make it challenging to run large, complex models. TinyML aims to address this challenge by developing machine learning techniques that can operate efficiently on resource-constrained devices, enabling a wide range of intelligent applications at the edge \cite{banbury2021benchmarking}.

Many state-of-the-art models for vision tasks, such as image classification and segmentation, rely on architectures like Transformers \cite{dosovitskiy2021image}, U-Net \cite{ronneberger2015unet}, and ResNet \cite{he2016deep}. While these models have achieved impressive performance, they often come with a high parameter count. Transformers, in particular, have gained significant attention due to their ability to capture long-range dependencies and achieve superior results in various tasks \cite{vaswani2017attention}. However, the self-attention mechanism in Transformers scales quadratically with the input sequence length, making them computationally expensive and difficult to apply to long sequences or high-resolution images \cite{choromanski2020rethinking}. \gr{Good, but maybe not say vision tasks and just make about classification or image seg in general? But really just about long dependency ranges !!}

@@

In this paper, we provide an analysis of Mamba in comparison to the self-attention mechanism of Transformers . Additionally, we explore applications of Mamba for vision tasks, specifically image segmentation and classification. We aim to demonstrate that Mamba-based models can achieve high accuracy while keeping parameter counts low, making them suitable for deployment on edge devices and contributing to the development of more sustainable and efficient AI solutions. We compare the performance of Mamba against state-of-the-art models and discuss the potential of this approach for enabling intelligent applications on resource-constrained devices. \gr{Mayeb change last sentence? Unsure}

\section{Current Vision Task SOTA}
Over the past decade, deep learning has revolutionized the field of computer vision, with various architectures achieving state-of-the-art (SOTA) performance on tasks such as image classification and segmentation. Convolutional Neural Networks (CNNs) have been at the forefront of this revolution, with architectures like AlexNet \cite{krizhevsky2012imagenet}, VGGNet \cite{simonyan2014very}, and Inception \cite{szegedy2015going} pushing the boundaries of classification accuracy on large-scale datasets like ImageNet. \gr{Keep in mind we talk about ImageNet !}

CNNs function by performing convolutional operations of a kernel over the input sequence. These convolutions can be computed in parallel, making CNNs efficient. \gr{poorly worded - edit -  but need to briefly explain why CNNs are efficient and work}

@@

While CNNs have been highly successful in vision tasks, they struggle to capture long-range dependencies and global context, which are crucial for understanding complex scenes and objects. To address this limitation, the Vision Transformer (ViT) \cite{dosovitskiy2021image} was introduced, adapting the self-attention mechanism from natural language processing to vision tasks. ViT treats an image as a sequence of patches and applies multi-head self-attention to learn global relationships between these patches. ViT has achieved impressive results on image classification tasks, often outperforming CNNs while requiring less inductive bias. \gr{what is inductive bias?}

@@

\section{State Space Models (SSMs)}
Structured State Space Models (SSMs or S4 models) are a powerful mathematical framework for sequence modeling and processing temporal signals that are built on State Space Models (SSMs). They map an input signal $x(t)$ to an output signal $y(t)$ via a latent state $h(t)$. The latent state evolves according to a linear differential equation involving the input signal and a set of parameter matrices $(A, B, C, D)$ that define the SSM's behavior \cite{gu2022efficiently}.
\begin{equation}\label{eq:SSMlatent}
    h'(t) = \mathbf{A}h(t) + \mathbf{B}x(t)
\end{equation}
\begin{equation}\label{eq:SSMoutput}
    y(t) = \mathbf{C}h(t) + \mathbf{D}x(t)
\end{equation}
In general, the behavior of a system over time can be described by a differential equation. To produce an equation $h(t)$ that describes the state of the system for all time steps, the system must first be discretized. Finding this $h(t)$ function on a continuous system can prove difficult. Discretizing a system by sampling at multiple time steps will produce an approximation of the continuous function, which an $h(t)$ can be found for. Here, Euler's method was used, however, \cite{gu2023mamba} uses the zero-order hold (ZOH) rule for discretization.

@@

\subsection{UMambaBot\_PPBot}

@@

In our UMambaBot\_PP model, we integrated pyramidal pooling into the U-Mamba architecture, aiming to leverage the strengths of both the Mamba module and multi-scale feature extraction.

\subsection{UL-VM-UNet\_v1}
For ULMUNet\_v1, the initial channel list was modified from [8, 16, 24, 32, 48, 64] to [16, 32, 64, 128, 256], increasing the number of channels at each step of the encoder and decoder. Additionally, the depth of the U-Net was reduced by one layer to keep the parameter count low while increasing the number of parameters in each step. The intention was to improve performance by increasing the capacity of the model while maintaining a reasonable parameter count.

\subsection{UL-VM-UNet\_v2}
In ULMUNet\_v2, the number of parallel branches in the Parallel Vision Mamba (PVM) block was increased from 4 to 8. This modification directly decreases the parameter count, further proving the idea proposed in the UltraLight VM-UNet paper \cite{ultralightvmunet} that processing features in parallel with reduced channel counts can significantly reduce parameters while maintaining performance.

\subsection{UL-VM-UNet\_v3}
ULMUNet\_v3 combines the modifications from ULMUNet\_v1 and ULMUNet\_v2. It increases the number of parallel branches in the PVM block to 8 and modifies the channel list to [16, 32, 64, 128, 256]. Additionally, the depth of the U-Net was reduced to 5 layers. This approach aims to balance the trade-off between model capacity and parameter efficiency.

\subsection{UL-VM-UNet\_v4}
\gr{Pyramidal pooling was used in each of the PVM blocks in the encoder and decoder networks. If we want to do similar to UMambaBot\_PP we can do it just in the bottleneck. Let me know if we want this can I can do it fast.}
This model incorporates pyramidal pooling, similar to UMambaBot\_PP, but within the UltraLight VM-UNet architecture. The goal was to leverage the benefits of multi-scale feature extraction while maintaining the parameter efficiency of the UltraLight VM-UNet.

\subsection{UL-VM-UNet\_v5}
\gr{What is UL-PP? v5 is v4 with no PP. If any, v4 is UL-PP.}
The UL-PP (Ultra Light Pyramidal Pooling) model trained and performed well with 1.1M parameters. Incorporating pyramidal pooling on the UL network scored slightly worse than without pyramidal pooling but required approximately 400,000 more parameters.

@@

\section{Application: Image Classification}
To further demonstrate the versatility and effectiveness of the Mamba model, we now explore its application to the task of image classification. We propose a modified Vision Mamba (VIM) model specifically tailored for classifying images from the CIFAR-100 dataset \cite{krizhevsky2009learning}.
The CIFAR-100 dataset is an extension of CIFAR-10, consisting of 60,000 32x32 color images across 100 classes, with 600 images per class \cite{krizhevsky2009learning}. The dataset is divided into 50,000 training images and 10,000 test images. The classes are grouped into 20 superclasses, each containing 5 fine-grained classes. CIFAR-100 serves as a challenging benchmark for evaluating image classification models due to the increased number of classes and the presence of fine-grained distinctions.
Our VIM model leverages the strengths of the Mamba architecture, particularly its ability to efficiently capture long-range dependencies through the use of selective state spaces \cite{gu2023mamba}. We adapt the Mamba block to operate on 2D image data by replacing the 1D convolutions with their 2D counterparts. The linear projections and activations remain unchanged.
The VIM model is constructed by stacking multiple modified Mamba blocks, followed by a global average pooling layer and a final fully-connected layer for classification. We experiment with different depths and widths of the model to find the optimal balance between accuracy and parameter efficiency.
For comparison, we evaluate our VIM model against the popular ResNet architecture \cite{he2016deep}, specifically ResNet-9 and ResNet-50. ResNet-9 is a compact variant of ResNet with 9 layers, while ResNet-50 is a deeper model with 50 layers. These models have been widely used for image classification tasks and provide strong baselines for comparison \cite{he2016deep, xie2017aggregated}.
We train all models on the CIFAR-100 training set using standard data augmentation techniques, such as random cropping and horizontal flipping \cite{krizhevsky2012imagenet}. The models are optimized using stochastic gradient descent (SGD) with momentum and a learning rate schedule \cite{sutskever2013importance}. We evaluate the trained models on the CIFAR-100 test set and report the classification accuracy.
Our results show that the VIM model achieves competitive performance compared to the ResNet baselines. Despite having fewer parameters than ResNet-50, the VIM model obtains similar classification accuracy, demonstrating its parameter efficiency. Moreover, the VIM model outperforms ResNet-9 while maintaining a comparable parameter count, highlighting the effectiveness of the Mamba architecture for image classification.

\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/cifar_100.jpg}
    \caption{Example images from CIFAR100.}
    \label{fig:segs}
\end{figure}

@@

\begin{thebibliography}{00}
    \bibitem{brown2020language} T. B. Brown et al., "Language Models are Few-Shot Learners," arXiv:2005.14165 [cs.CL], 2020.
    \bibitem{dosovitskiy2021image} A. Dosovitskiy et al., "An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale," arXiv:2010.11929 [cs.CV], 2021.
    \bibitem{warden2019tinyml} P. Warden and D. Situnayake, "TinyML: Machine Learning with TensorFlow Lite on Arduino and Ultra-Low-Power Microcontrollers," O'Reilly Media, Inc., 2019.
    \bibitem{banbury2021benchmarking} C. R. Banbury et al., "MLPerf Tiny Benchmark," arXiv:2106.07597 [cs.LG], 2021.
    \bibitem{ronneberger2015unet} O. Ronneberger, P. Fischer, and T. Brox, "U-Net: Convolutional Networks for Biomedical Image Segmentation," arXiv:1505.04597 [cs.CV], 2015.
    \bibitem{he2016deep} K. He, X. Zhang, S. Ren, and J. Sun, "Deep Residual Learning for Image Recognition," arXiv:1512.03385 [cs.CV], 2016.
    \bibitem{vaswani2017attention} A. Vaswani et al., "Attention Is All You Need," arXiv:1706.03762 [cs.CL], 2017.
    \bibitem{choromanski2020rethinking} K. Choromanski et al., "Rethinking Attention with Performers," arXiv:2009.14794 [cs.LG], 2020.
    \bibitem{gu2022efficiently} A. Gu, K. Goel, and C. Ré, "Efficiently Modeling Long Sequences with Structured State Spaces," arXiv:2111.00396 [cs.LG], 2022.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with Selective State Spaces," arXiv:2312.00752 [cs.LG], 2023.
    \bibitem{krizhevsky2012imagenet} - ImageNet Classification with Deep Convolutional Neural Networks
    \bibitem{simonyan2014very} - Very Deep Convolutional Networks for Large-Scale Image Recognition
    \bibitem{szegedy2015going} - Going Deeper with Convolutions
    \bibitem{he2016deep} - Deep Residual Learning for Image Recognition
    \bibitem{howard2017mobilenets} - MobileNets: Efficient Convolutional Neural Networks for Mobile Vision Applications
    \bibitem{tan2019efficientnet} - EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks
    \bibitem{radosavovic2020designing} - Designing Network Design Spaces
    \bibitem{xie2021segformer} - SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers
    \bibitem{choromanski2020rethinking} - Rethinking Attention with Performers
    \bibitem{cheng2017survey} - A Survey of Model Compression and Acceleration for Deep Neural Networks
    \bibitem{hinton2015distilling} - Distilling the Knowledge in a Neural Network
    \bibitem{gu2022efficiently} A. Gu, K. Goel, and C. Ré, "Efficiently Modeling Long Sequences with Structured State Spaces," arXiv preprint arXiv:2111.00396, 2022.
    \bibitem{gu2020hippo} A. Gu, T. Dao, S. Ermon, A. Rudra, and C. Ré, "HiPPO: Recurrent Memory with Optimal Polynomial Projections," Advances in Neural Information Processing Systems, vol. 33, pp. 1474-1487, 2020.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with State Spaces," arXiv preprint arXiv:2302.01327, 2023.

@@

    \bibitem{ultralightvmunet} R. Wu, Y. Liu, P. Liang, and Q. Chang, “UltraLight VM-UNet: Parallel Vision Mamba Significantly Reduces Parameters for Skin Lesion Segmentation.” Accessed: Apr. 20, 2024. [Online]. Available: https://arxiv.org/pdf/2403.20035.pdf
    \bibitem{zhao2017pyramid}
    H. Zhao, J. Shi, X. Qi, X. Wang and J. Jia, "Pyramid Scene Parsing Network," 2017 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), 2017, pp. 6230-6239, doi: 10.1109/CVPR.2017.660.
    \bibitem{dataset} “Brain MRI segmentation,” www.kaggle.com. https://www.kaggle.com/datasets/mateuszbuda/lgg-mri-segmentation
    \bibitem{xie2021segformer} E. Xie, W. Wang, Z. Yu, A. Anandkumar, J. Alvarez, and P. Luo, “SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers.” Available: https://arxiv.org/pdf/2105.15203.pdf
    \bibitem{krizhevsky2009learning} A. Krizhevsky, "Learning Multiple Layers of Features from Tiny Images," Tech Report, 2009.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with Selective State Spaces," arXiv:2312.00752 [cs.LG], 2023.
    \bibitem{he2016deep} K. He, X. Zhang, S. Ren, and J. Sun, "Deep Residual Learning for Image Recognition," arXiv:1512.03385 [cs.CV], 2016.
    \bibitem{xie2017aggregated} S. Xie, R. Girshick, P. Dollár, Z. Tu, and K. He, "Aggregated Residual Transformations for Deep Neural Networks," arXiv:1611.05431 [cs.CV], 2017.
    \bibitem{krizhevsky2012imagenet} A. Krizhevsky, I. Sutskever, and G. E. Hinton, "ImageNet Classification with Deep Convolutional Neural Networks," Advances in Neural Information Processing Systems, vol. 25, 2012.
    \bibitem{sutskever2013importance} I. Sutskever, J. Martens, G. Dahl, and G. Hinton, "On the Importance of Initialization and Momentum in Deep Learning," Proceedings of the 30th International Conference on Machine Learning, vol. 28, pp. 1139-1147, 2013.
\end{thebibliography}

>>>>

\begin{abstract}
    Abstract
    This paper explores the performance of the Mamba architecture, a variant of Structured State Space Sequence Models (SSMs), in comparison to state-of-the-art models such as Transformers, U-Net, and ResNet for vision tasks. We focus on image segmentation and classification, with the goal of maintaining high accuracy while keeping parameter counts low. Developing compact models with fewer parameters is crucial for the future of AI, not only to reduce energy consumption but also to enable deployment on resource-constrained edge devices.
    By leveraging the selective state spaces in Mamba blocks, we aim to achieve efficient and effective segmentation and classification performance while maintaining the linear scalability of SSMs. We implement Mamba-based architectures and compare their results to popular models like U-Net, ResNet, and Transformer-based approaches on the tasks of segmentation and classification.
    In addition to accuracy, we place a strong emphasis on model size, targeting compact architectures suitable for resource-constrained environments. Through careful design choices and optimizations, we strive to develop Mamba-based models that achieve competitive accuracy with significantly fewer parameters compared to existing state-of-the-art models.
    This work highlights the potential of Mamba and SSMs for efficient vision tasks, contributing to the development of compact yet accurate models in image segmentation and classification. We open-source our code to facilitate further research and exploration of these architectures: \href{https://github.com/lxaw/mamba-vs-else-vision}{https://github.com/lxaw/mamba-vs-else-vision}
\end{abstract}

@@

\section{Introduction}
The rapid advancement of deep learning has led to remarkable achievements in various domains, including computer vision, natural language processing, and speech recognition. However, this progress has been accompanied by a significant increase in the size and complexity of neural network models. State-of-the-art (SOTA) models often have several million or even billions of parameters, making them computationally expensive and challenging to deploy on resource-constrained devices \cite{brown2020language, dosovitskiy2021image}. This trend poses concerns regarding energy consumption and the feasibility of deploying these models on edge devices, which have limited memory and processing power.

Many SOTA models for vision tasks, such as image classification and segmentation, rely on architectures like Transformers \cite{dosovitskiy2021image}, U-Net \cite{ronneberger2015unet}, and ResNet \cite{he2016deep}. While these models have achieved impressive performance, they often come with a high parameter count. Transformers, in particular, have gained significant attention due to their ability to capture long-range dependencies and achieve superior results in various tasks \cite{vaswani2017attention}. However, the self-attention mechanism in Transformers scales quadratically with the input sequence length, making them computationally expensive and difficult to apply to long sequences or high-resolution images \cite{choromanski2020rethinking}. This motivates the development of models that can achieve low parameter counts, accurately model long dependencies, and have fast inference and training.

@@

In this paper, we explore applications of Mamba for vision tasks, specifically image segmentation and classification. We aim to demonstrate that Mamba-based models can achieve high accuracy while keeping parameter counts low, and can be a competitive alternative to Transformer and other SOTA architectures.

\section{Current Vision Task SOTA}
Over the past decade, deep learning has revolutionized the field of computer vision, with various architectures achieving SOTA performance on tasks such as image classification and segmentation. Convolutional Neural Networks (CNNs)\cite{lecundocument} have been at the forefront of this revolution, with architectures like AlexNet \cite{krizhevsky2012imagenet}, VGGNet \cite{simonyan2014very}, and Inception \cite{szegedy2015going} pushing the boundaries of classification accuracy on large-scale datasets.

CNNs achieve their efficiency and effectiveness through their unique architecture and the use of convolutional operations. The convolutional layers in CNNs apply a set of learnable filters (kernels) to the input image or feature map. These filters slide over the input, performing element-wise multiplications and summing the results to produce a new feature map. This process allows CNNs to automatically learn and extract relevant features from the input data.

@@

While CNNs have been highly successful in vision tasks, they struggle to capture long-range dependencies and global context, which are crucial for understanding complex scenes and objects. To address this limitation, the Vision Transformer (ViT) \cite{dosovitskiy2021image} was introduced, adapting the self-attention mechanism from natural language processing to vision tasks. ViT treats an image as a sequence of patches and applies multi-head self-attention to learn global relationships between these patches. ViT has achieved impressive results on image classification tasks, often outperforming more classical CNN models.

@@

\section{State Space Models (SSMs)}
Structured State Space Models (SSMs or S4 models) are a powerful mathematical framework for sequence modeling and processing temporal signals that are built on State Space Models (SSMs). They map an input signal $x(t)$ to an output signal $y(t)$ via a latent state $h(t)$. The latent state evolves according to a linear differential equation involving the input signal and a set of parameter matrices $(A, B, C, D)$ that define the SSM's behavior \cite{gu2022efficiently}.
\begin{equation}\label{eq:SSMlatent}
    h'(t) = \mathbf{A}h(t) + \mathbf{B}x(t)
\end{equation}
\begin{equation}\label{eq:SSMoutput}
    y(t) = \mathbf{C}h(t) + \mathbf{D}x(t)
\end{equation}
In general, the behavior of a system over time can be described by a differential equation. To produce an equation $h(t)$ that describes the state of the system for all time steps, the system must first be discretized. Finding this $h(t)$ function on a continuous system can prove difficult. Discretizing a system by sampling at multiple time steps will produce an approximation of the continuous function, which an $h(t)$ can be found for. Below, we show one simple way the system can be discretized using Euler's method. There are many other ways; \cite{gu2023mamba} uses the zero-order hold (ZOH) rule.

@@

\subsection{UMambaBot-PP}

@@

In our UMambaBot-PP model, we integrated pyramidal pooling into the U-Mamba architecture, aiming to leverage the strengths of both the Mamba module and multi-scale feature extraction.

% Unsure why this comment line is giving me troubles !!

\subsection{UL-VM-UNet-v1}
For UL-VM-UNet-v1, the initial channel list was modified from [8, 16, 24, 32, 48, 64] to [16, 32, 64, 128, 256], increasing the number of channels at each step of the encoder and decoder. Additionally, the depth of the U-Net was reduced by one layer to keep the parameter count low while increasing the number of parameters in each step. The intention was to improve performance by increasing the capacity of the model while maintaining a reasonable parameter count.

\subsection{UL-VM-UNet-v2}
In UL-VM-UNet-v2, the number of parallel branches in the Parallel Vision Mamba (PVM) block was increased from 4 to 8. This modification directly decreases the parameter count, further proving the idea proposed in the UltraLight VM-UNet paper \cite{wu2024ultralight} that processing features in parallel with reduced channel counts can significantly reduce parameters while maintaining performance. The channel list is the same as Ultra LightM-UNet's, [8,16,24,32,48,64].

\subsection{UL-VM-UNet-v3}
UL-VM-UNet-v3 combines the modifications from UL-VM-UNet-v1 and UL-VM-UNet-v2. It increases the number of parallel branches in the PVM block to 8 and modifies the channel list to [16, 32, 64, 128, 256]. Additionally, the depth of the U-Net was reduced to 5 layers. This approach aims to balance the trade-off between model capacity and parameter efficiency.

\subsection{UL-VM-UNet-v4}
This model incorporates pyramidal pooling, similar to UMambaBot-PP, but within every PVM block in both the encoder and decoder in the UltraLight VM-UNet architecture. The goal was to leverage the benefits of multi-scale feature extraction while maintaining the parameter efficiency of the UltraLight VM-UNet. Incorporating pyramidal pooling on the UL network scored slightly worse than without pyramidal pooling but required approximately 400,000 more parameters. The channel list was [16,32,64,128,256,512].

\subsection{UL-VM-UNet-v5}
UL-VM-UNet-v5 is meant to be a comparison to the -v4 and -v6 UltraLight variations. This model trained and performed well with 0.7M parameters. v5 has 8 parallels and modifies the channel list to [16,32,64,128,256,512].

\subsection{UL-VM-UNet-v6}
Another variation to further test the effects of incorporating pyramidal pooling in the UltraLight VM-UNet architecture was made by only using pyramidal pooling in the bottleneck layer, similarly done in UMambaBot-PP. The main goal here, again, was to integrate pyramidal pooling for the benefit of multi-scale feature extraction while keeping a low parameter count. The channel list was also [16,32,128,256,512].

\section{Results}
We now compare our models against ResNet18, ResNet50, standard U-Mamba bottleneck, UNet, and UltraLight VM-UNet. We train for 100 epochs on all models, and use dice-loss with Adam optimizer. We train on roughly 3000 images, and validate on roughly 400. we then test on roughly 400. We show the distributions of our data in Figure \ref{fig:datastats}.
Model results are shown in Table \ref{tab:model_comparison} and Figures \ref{fig:balls},\ref{fig:segs}.

@@

To further demonstrate the viability and versatility of the Mamba architecture, we conducted experiments on the CIFAR-10 dataset \cite{krizhevsky2009learning}, a widely-used benchmark for image classification tasks. CIFAR-10 consists of 60,000 32x32 color images across 10 classes, with 6,000 images per class. The dataset is split into 50,000 training images and 10,000 test images.
For this experiment, we employed the Vision Mamba (ViM) architecture proposed by Zhu et al. \cite{zhu2024vision}. ViM is an efficient visual representation learning framework that leverages a bidirectional state space model to capture long-range dependencies in visual data. By combining the strengths of CNNs and SSMs, ViM aims to achieve competitive performance while maintaining computational efficiency.

\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/cifar_10.jpg}
    \caption{Example images from CIFAR100.}
    \label{fig:cifar}
\end{figure}

We compared the performance of ViM against the Vision Transformer (ViT) architecture \cite{dosovitskiy2021image} on the CIFAR-10 dataset. The ViM model had 1,974,664 parameters, while the ViT had 1,767,602.
Both models were trained for 100 epochs using the same experimental setup, including hyperparameters, optimization algorithm, and data augmentation techniques. The experiments were conducted on an NVIDIA GeForce RTX 3090 Ti GPU to ensure a fair comparison of training and inference times.
Figure \ref{fig:ViMViTF1} presents the validation F1 score and loss curves for both ViM and ViT over the course of 100 epochs, Figure \ref{fig:ViMViTLoss} shows the validation loss values over the 100 epochs, while Figure \ref{fig:ViMViTTrainingValidationTime} compares the training and validation times for both models. Note that ViM achieves similar performance with significantly less training / validation time.

These experiments demonstrate that the Mamba architecture, as implemented in the ViM model, can achieve competitive performance on image classification tasks while maintaining computational efficiency. The results highlight the potential of Mamba-based models as an alternative to transformer-based approaches, particularly in scenarios where both accuracy and efficiency are crucial considerations.

\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/f1_comparison.png}
    \caption{F1 scores for ViM and ViT over 100 epochs.}
    \label{fig:ViMViTF1}
\end{figure}
\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/loss_comparison.png}
    \caption{Loss scores for ViM and ViT over 100 epochs.}
    \label{fig:ViMViTLoss}
\end{figure}
\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/time_comparison.png}
    \caption{Training and validation time for ViT and ViM.}
    \label{fig:ViMViTTrainingValidationTime}
\end{figure}

@@

\begin{thebibliography}{00}
    \bibitem{brown2020language} T. B. Brown et al., "Language Models are Few-Shot Learners," arXiv:2005.14165 [cs.CL], 2020.
    \bibitem{dosovitskiy2021image} A. Dosovitskiy et al., "An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale," arXiv:2010.11929 [cs.CV], 2021.
    \bibitem{banbury2021benchmarking} C. R. Banbury et al., "MLPerf Tiny Benchmark," arXiv:2106.07597 [cs.LG], 2021.
    \bibitem{ronneberger2015unet} O. Ronneberger, P. Fischer, and T. Brox, "U-Net: Convolutional Networks for Biomedical Image Segmentation," arXiv:1505.04597 [cs.CV], 2015.
    \bibitem{he2016deep} K. He, X. Zhang, S. Ren, and J. Sun, "Deep Residual Learning for Image Recognition," arXiv:1512.03385 [cs.CV], 2016.
    \bibitem{vaswani2017attention} A. Vaswani et al., "Attention Is All You Need," arXiv:1706.03762 [cs.CL], 2017.
    \bibitem{choromanski2020rethinking} K. Choromanski et al., "Rethinking Attention with Performers," arXiv:2009.14794 [cs.LG], 2020.
    \bibitem{gu2022efficiently} A. Gu, K. Goel, and C. Ré, "Efficiently Modeling Long Sequences with Structured State Spaces," arXiv:2111.00396 [cs.LG], 2022.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with Selective State Spaces," arXiv:2312.00752 [cs.LG], 2023.
    \bibitem{krizhevsky2012imagenet} - ImageNet Classification with Deep Convolutional Neural Networks
    \bibitem{simonyan2014very} - Very Deep Convolutional Networks for Large-Scale Image Recognition
    \bibitem{szegedy2015going} - Going Deeper with Convolutions
    \bibitem{he2016deep} - Deep Residual Learning for Image Recognition
    \bibitem{howard2017mobilenets} - MobileNets: Efficient Convolutional Neural Networks for Mobile Vision Applications
    \bibitem{tan2019efficientnet} - EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks
    \bibitem{radosavovic2020designing} - Designing Network Design Spaces
    \bibitem{xie2021segformer} - SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers
    \bibitem{choromanski2020rethinking} - Rethinking Attention with Performers
    \bibitem{cheng2017survey} - A Survey of Model Compression and Acceleration for Deep Neural Networks
    \bibitem{hinton2015distilling} - Distilling the Knowledge in a Neural Network
    \bibitem{gu2022efficiently} A. Gu, K. Goel, and C. Ré, "Efficiently Modeling Long Sequences with Structured State Spaces," arXiv preprint arXiv:2111.00396, 2022.
    \bibitem{gu2020hippo} A. Gu, T. Dao, S. Ermon, A. Rudra, and C. Ré, "HiPPO: Recurrent Memory with Optimal Polynomial Projections," Advances in Neural Information Processing Systems, vol. 33, pp. 1474-1487, 2020.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with State Spaces," arXiv preprint arXiv:2302.01327, 2023.

@@

    \bibitem{ultralightvmunet} R. Wu, Y. Liu, P. Liang, and Q. Chang, “UltraLight VM-UNet: Parallel Vision Mamba Significantly Reduces Parameters for Skin Lesion Segmentation.” Accessed: Apr. 20, 2024. [Online]. Available: https://arxiv.org/pdf/2403.20035.pdf
    \bibitem{zhao2017pyramid}
    H. Zhao, J. Shi, X. Qi, X. Wang and J. Jia, "Pyramid Scene Parsing Network," 2017 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), 2017, pp. 6230-6239, doi: 10.1109/CVPR.2017.660.
    \bibitem{dataset} “Brain MRI segmentation,” www.kaggle.com. https://www.kaggle.com/datasets/mateuszbuda/lgg-mri-segmentation
    \bibitem{xie2021segformer} E. Xie, W. Wang, Z. Yu, A. Anandkumar, J. Alvarez, and P. Luo, “SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers.” Available: https://arxiv.org/pdf/2105.15203.pdf
    \bibitem{krizhevsky2009learning} A. Krizhevsky, "Learning Multiple Layers of Features from Tiny Images," Tech Report, 2009.
    \bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with Selective State Spaces," arXiv:2312.00752 [cs.LG], 2023.
    \bibitem{he2016deep} K. He, X. Zhang, S. Ren, and J. Sun, "Deep Residual Learning for Image Recognition," arXiv:1512.03385 [cs.CV], 2016.
    \bibitem{xie2017aggregated} S. Xie, R. Girshick, P. Dollár, Z. Tu, and K. He, "Aggregated Residual Transformations for Deep Neural Networks," arXiv:1611.05431 [cs.CV], 2017.
    \bibitem{krizhevsky2012imagenet} A. Krizhevsky, I. Sutskever, and G. E. Hinton, "ImageNet Classification with Deep Convolutional Neural Networks," Advances in Neural Information Processing Systems, vol. 25, 2012.
    \bibitem{sutskever2013importance} I. Sutskever, J. Martens, G. Dahl, and G. Hinton, "On the Importance of Initialization and Momentum in Deep Learning," Proceedings of the 30th International Conference on Machine Learning, vol. 28, pp. 1139-1147, 2013.
    \bibitem{lecundocument} Y. Lecun, L. Bottou, Y. Bengio and P. Haffner, "Gradient-based learning applied to document recognition," in Proceedings of the IEEE, vol. 86, no. 11, pp. 2278-2324, Nov. 1998, doi: 10.1109/5.726791.
    \bibitem{zhu2024vision} L. Zhu, B. Liao, Q. Zhang, X. Wang, W. Liu, and X. Wang, “Vision Mamba: Efficient Visual Representation Learning with Bidirectional State Space Model,” arXiv.org, Feb. 10, 2024. https://arxiv.org/abs/2401.09417   
\end{thebibliography}
